The Power of Two Random Choices

### 2014-03-31 presentation/presentation.tex
group references by importance

\documentclass[serif,professionalfonts]{beamer}

\usetheme{Forcellini}

@@

%---------------------------------------------------------------------
%Front matter
%---------------------------------------------------------------------
\title{The Power of Two Random Choices}
\subtitle{And the Surprising Influence of Asymmetry}
\author[M. Kalany]{Martin Kalany}
\institute[TU Wien]
{
  Graduate student in Computer Science\\
  Vienna University of Technology\\
}
\date{\today}

@@

\medskip
\begin{theorem}[Raab and Steger, 1998]
When using the single choice strategy to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\loadS \leq 
	\begin{cases}
    \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases} \quad .
\end{align*}
\end{theorem}

@@

\bigskip
We can distinguish placement algorithms by the way the $d$ \alert{bins are sampled}:
\begin{enumerate}
\item Uniformly and independently
\item Non-uniformly and independently
\item Non-uniformly and dependently
\end{enumerate}
\end{frame}

@@

\begin{frame}
\frametitle{Single choice vs.~Greedy}
\begin{align*}
\loadS \leq 
	\begin{cases}
    \frac{\ln n}{\ln\ln n} + O(1)              & m = n \\
    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases}
\end{align*}
\begin{align*}
\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) 
\end{align*}

@@

$d$-step Fibonacci numbers:
\begin{align*}
F_d(k) := \begin{cases}
                0               & k \leq 0\\
                1               & k = 1\\
                \sum_{i=1}^{d}F_d(k-i) & k > 1
            \end{cases}
\end{align*}
Generalized golden ratio $\Phi_d$:
\begin{align*}
\Phi_d := \lim_{k \rightarrow \infty} \frac{F_d(k)}{F_d(k-1)}
\end{align*}
\begin{align*}
\Phi_2 = 1.61 \quad , \quad \forall i\; \forall j \in \mathbb{N}: i < j \implies \Phi_i < \Phi_j \quad , \quad \lim_{d\rightarrow \infty} \Phi_d = 2 \quad .
\end{align*}
\begin{align*}
F_d(k) \geq \Phi_d^{k-2}
\end{align*}
\end{frame}

@@

\begin{frame}
\frametitle{The Always-go-left algorithm}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{align*}
\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) 
\end{align*}
\end{column}
\begin{column}{0.5\textwidth}
\begin{align*}
\loadAgl \leq \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
\end{align*}
\end{column}
\end{columns}
\bigskip
The combination of \alert{asymmetric tie breaking} and \alert{non-uniform} (and independent) sampling of the $d$ bins is crucial to achieve this upper bound.
\bigskip
\begin{theorem}[V\"ocking, 2003]
For $m=n$, using an arbitrary allocation scheme that chooses $d$ bins for each ball at random (i.e., class 3) the maximum load is, w.h.p,
\begin{align*}
\loadAgl = \frac{\ln \ln n}{d \ln \Phi_d} - O(1) \quad .
\end{align*}
\end{theorem}
\end{frame}

@@

\begin{frame}
\frametitle{Proof of activation}
\begin{block}{Lemma}
The occurrence of a bad event (i.e., a bin contains more than $L+3$) implies the activation of an AWT of order $L$.
\end{block}
\bigskip
\visible<2>{
We assign balls to the nodes of the AWT s.t.:
\myInvariant
}
\end{frame}

@@

}
\end{enumerate}

@@

\begin{frame}
\frametitle{Edge events}
\defEdgeEvent
\bigskip
One edge is activated with probability at most $d/n$ and we have $k-1$ edges; thus
\begin{align*}
p_\mathrm{e} = \left(\frac{d}{n}\right)^{k-1} \quad .
\end{align*}
\end{frame}

@@

\begin{frame}
\frametitle{Probability of activation}
Putting it together, we get
\begin{align*}
P_\mathrm{a} = \omega p_e p_l \leq  n^k \left(\frac{d}{n}\right)^{k-1} 3^{-dl} \quad .
\end{align*}
Using $ k\leq 2l$, $2d^2\leq 3^d$ and $l \geq \Phi_d^{dL-1}$, we can simplify to
\begin{align*}
P_\mathrm{a} \leq n 2 ^{-\Phi_d^{dL-1}} \quad .
\end{align*}
For arbitrary $\alpha > 0$, let
\begin{align*}
L &\geq \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}\right\rceil+\frac{1}{d} \implies \frac{\ln\ln n}{d \ln \Phi_d} + O\left(1\right) ,
\end{align*}
and we get
\begin{align*}
P_\mathrm{a} \leq n^{-\alpha} \xqed{5cm}
\end{align*}
\end{frame}

@@

\begin{frame}
\frametitle{Conclusion}
\begin{enumerate}
\item The balls-into-bins model \alert{adopts naturally} to various problems (hashing, load balancing,\dots).
\item Applying the two-choice paradigm reduces the upper bound for the maximum load by an \alert{exponential factor} (compared to a single choice scheme).
\item \alert{Many variations} possible: Non-uniform balls/bins; weighted balls; parallel allocation of several balls;\dots
\item \alert{Witness trees} are a suitable proof technique for this kind of problems.
\end{enumerate}
\end{frame}

\begin{frame}[shrink]
\frametitle<presentation>{References}    
\bibliographystyle{acm}
\nocite{VOC03}
\nocite{ABKU99}
\nocite{RS98}
\nocite{BCSV06}
\nocite{capocelli1990generalized}
\setbeamertemplate{bibliography item}[text]
\bibliography{../sources}
\end{frame}

>>>>

\documentclass[serif,professionalfonts]{beamer}
\usetheme{Forcellini}

\usepackage[backend=bibtex]{biblatex} 
\addbibresource{../sources.bib}

@@

%---------------------------------------------------------------------
%Front matter
%---------------------------------------------------------------------
\title{The Power of Two Random Choices}
\subtitle{\dots and the Surprising Influence of Asymmetry}
\author[M. Kalany]{Martin Kalany}
\institute[TU Wien]
{
  Graduate student in Computer Science\\
  Vienna University of Technology\\
}
\date{\today}

@@

\medskip
\begin{theorem}[Raab and Steger, 1998]
When using the \alert{single choice} strategy to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\loadS \leq 
	\begin{cases}
    \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases} \quad .
\end{align*}
\end{theorem}
\end{frame}

@@

\bigskip
We distinguish placement algorithms by the way the $d$ \alert{bins are sampled}:
\begin{enumerate}
\item Uniformly and independently
\item Non-uniformly and independently
\item Non-uniformly and dependently
\end{enumerate}
\end{frame}

@@

\begin{frame}
\frametitle{Single choice vs.~Greedy}
\begin{align*}
\loadS \leq 
	\begin{cases}
    \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases}
\end{align*}
\begin{align*}
\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) 
\end{align*}

@@

$d$-step Fibonacci numbers:
\begin{align*}
F_d(k) := \begin{cases}
                0               & k \leq 0\\
                1               & k = 1\\
                \sum_{i=1}^{d}F_d(k-i) & k > 1
            \end{cases}
\end{align*}
Generalized golden ratio $\Phi_d$:
\begin{align*}
\Phi_d := \lim_{k \rightarrow \infty} \frac{F_d(k)}{F_d(k-1)}
\end{align*}
\begin{align*}
\Phi_2 = 1.61 \quad , \quad \forall i\; \forall j \in \mathbb{N}: i < j \implies \Phi_i < \Phi_j \quad , \quad \lim_{d\rightarrow \infty} \Phi_d = 2 
\end{align*}
\begin{align*}
F_d(k) \geq \Phi_d^{k-2}
\end{align*}
\end{frame}

@@

\begin{frame}
\frametitle{The Always-go-left algorithm}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{align*}
\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) 
\end{align*}
\end{column}
\begin{column}{0.5\textwidth}
\begin{align*}
\loadAgl \leq \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
\end{align*}
\end{column}
\end{columns}
\bigskip
The combination of \alert{asymmetric tie breaking} and \alert{non-uniform} (and independent) sampling of the $d$ bins is crucial to achieve this upper bound.
\bigskip
\begin{theorem}[V\"ocking, 2003]
For $m=n$, using an arbitrary allocation scheme that chooses $d$ bins for each ball at random (i.e., class 3), the maximum load is, w.h.p,
\begin{align*}
\loadAgl = \frac{\ln \ln n}{d \ln \Phi_d} - O(1) \quad .
\end{align*}
\end{theorem}
\end{frame}

@@

\begin{frame}
\frametitle{Proof of activation}
\begin{block}{Lemma}
The occurrence of a bad event (i.e., a bin contains more than $L+3$ balls) implies the activation of an AWT of order $L$.
\end{block}
\bigskip
\visible<2>{
We assign balls to the nodes of the AWT s.t.:
\myInvariant
}
\end{frame}

@@

}
\end{enumerate}
\end{frame}

@@

\begin{frame}
\frametitle{Edge events}
\defEdgeEvent
\bigskip
One edge is activated with probability at most $d/n$ and we have $k-1$ edges; thus
\begin{align*}
p_\mathrm{e} \leq \left(\frac{d}{n}\right)^{k-1} \quad .
\end{align*}
\end{frame}

@@

\begin{frame}
\frametitle{Probability of activation}
Putting it together, we get
\begin{align*}
P_\mathrm{a} \leq \omega p_e p_l \leq  n^k \left(\frac{d}{n}\right)^{k-1} 3^{-dl} \quad .
\end{align*}
Using $ k\leq 2l$, $2d^2\leq 3^d$ and $l \geq \Phi_d^{dL-1}$, we can simplify to
\begin{align*}
P_\mathrm{a} \leq n 2 ^{-\Phi_d^{dL-1}} \quad .
\end{align*}
For arbitrary $\alpha > 0$, let
\begin{align*}
L &\geq \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}\right\rceil+\frac{1}{d} \implies \frac{\ln\ln n}{d \ln \Phi_d} + O\left(1\right) ,
\end{align*}
and we get
\begin{align*}
P_\mathrm{a} \leq n^{-\alpha} \xqed{5cm}
\end{align*}
\end{frame}

@@

\begin{frame}
\frametitle{Conclusion}
\begin{enumerate}
\item The balls-into-bins model \alert{adopts naturally} to various problems (hashing, load balancing,\dots).
\item Applying the two-choice paradigm reduces the upper bound for the maximum load by an \alert{exponential factor} (compared to the single choice scheme).
\item \alert{Many variations} possible: Non-uniform balls/bins; weighted balls; parallel allocation of several balls;\dots
\item \alert{Witness trees} are a suitable proof technique for this kind of problems.
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{References}
\nocite{VOC03}
\nocite{ABKU99}
\nocite{RS98}
\nocite{BCSV06}
\nocite{capocelli1990generalized}
\setbeamertemplate{bibliography item}[triangle]
\printbibliography[keyword=main]
\bigskip

\printbibliography[notkeyword=main]
\end{frame}
